Scratpad

```{r}
normalized_df2 = data.frame(matrix(ncol = length(colnames(train_df)), nrow = 0))

for (i in 1:nrow(train_df)) {
  temp = train_df[i, "temp"]
  bpSys = train_df[i, "bpSys"]
  vo2 = train_df[i, "vo2"]
  throat = train_df[i, "throat"]
  atRisk = train_df[i, "atRisk"]
  headA = train_df[i, "headA"]
  bodyA = train_df[i, "bodyA"]
  cough = train_df[i, "cough"]
  runny = train_df[i, "runny"]
  nausea = train_df[i, "nausea"]
  diarrhea = train_df[i, "diarrhea"]
  
  row <- c()
  
  for(col_name in colnames(train_df)) {
    value <- train_df[i, col_name]
    norm_value <- (value - min(train_df[col_name])) / (max(train_df[col_name]) - min(train_df[col_name]))
    
    row <- c(row, norm_value)
  }
  
  #v2temp = (temp - min(train_df["temp"])) / (max(train_df["temp"]) - min(train_df["temp"]))
  #v2bpSys = (bpSys - min(train_df["bpSys"])) / (max(train_df["bpSys"]) - min(train_df["bpSys"]))
  #v2vo2 = (vo2 - min(train_df["vo2"])) / (max(train_df["vo2"]) - min(train_df["vo2"]))
  #v2throat = (throat - min(train_df["throat"])) / (max(train_df["throat"]) - min(train_df["throat"]))
  #v2atRisk = (atRisk - min(train_df["atRisk"])) / (max(train_df["atRisk"]) - min(train_df["atRisk"]))
  #v2headA = (headA - min(train_df["headA"])) / (max(train_df["headA"]) - min(train_df["headA"]))
  #v2bodyA = (bodyA - min(train_df["bodyA"])) / (max(train_df["bodyA"]) - min(train_df["bodyA"]))
  #v2cough = (cough - min(train_df["cough"])) / (max(train_df["cough"]) - min(train_df["cough"]))
  #v2runny = (runny - min(train_df["runny"])) / (max(train_df["runny"]) - min(train_df["runny"]))
  #v2nausea = (nausea - min(train_df["nausea"])) / (max(train_df["nausea"]) - min(train_df["nausea"]))
  #v2diarrhea = (diarrhea - min(train_df["diarrhea"])) / (max(train_df["diarrhea"]) - min(train_df["diarrhea"]))
  
  #normalized_df[nrow(normalized_df) + 1,] = c(
   # v2temp,
   # v2bpSys,
   #  v2vo2,
   # v2throat,
  #  v2atRisk,
   # v2headA,
  #  v2bodyA,
   # v2cough,
  #  v2runny,
  #  v2nausea,
  #  v2diarrhea
  #)
  normalized_df2[nrow(normalized_df2) + 1, ] = row
}
```


```{r}
set.seed(10111)
x = matrix(rnorm(40), 20, 2)
y = rep(c(-1, 1), c(10, 10))
x[y == 1,] = x[y == 1,] + 1
dat = data.frame(x, y = as.factor(y))
dat
```
```{r}
#calculating the bin ranges
bin_ranges <- function(n, col, df) {
  my_vec <- c(df[, col])
  my_vec <- sort(my_vec)
  min <- min(my_vec)
  max <- max(my_vec)
  rnge <- (max - min) / n
  print(n)
  for (i in 1:n) {
    print(paste("min:", min, "max:", max, "rnge:", rnge))
    min <- floor(min + rnge)
    
    
  }
  vecs <- split(my_vec,             # Applying split() function
                cut(seq_along(my_vec),
                    n,
                    labels = FALSE))
  print(lengths(vecs))
  return (vecs)
  
}

rng<-bin_ranges(3,"throat",train_df)
```

```{r}
norm_values = list(
  temp = c(90, 106),
  bpSys = c(90, 150),
  vo2 = c(10, 70),
  throat = c(80, 120),
  atRisk = c(0, 1),
  headA = c(0, 9),
  bodyA = c(0, 9),
  cough = c(0, 1),
  runny = c(0, 1),
  nausea = c(0, 1),
  diarrhea = c(0, 1)
)

for(item in colnames(train_df)) {
  if (item %in% names(norm_values)) {
    ind = between(train_df[,item], norm_values[[item]][1], norm_values[[item]][2])
    print(paste(item, train_df[!ind, item], length(train_df[!ind, item])))
  }
}
```

```{r}

normalized_df = data.frame(matrix(ncol = length(colnames(train_df)), nrow = 0))
colnames(normalized_df) = colnames(train_df)

for (i in 1:nrow(train_df)) {
  row <- c()
  
  for (col_name in colnames(train_df)) {
    value <- train_df[i, col_name]
    norm_value <-
      (value - min(train_df[col_name])) / (max(train_df[col_name]) - min(train_df[col_name]))
    
    row <- c(row, norm_value)
  }
  
  normalized_df[nrow(normalized_df) + 1,] = row
}

head(normalized_df)
```


```{r}
dat <- matrix(data=runif(n=4, min=0, max=1), nrow=2, ncol=2, 
              dimnames=list(c("pos", "neg"), c("pos", "neg")))

# now set the names *of the dimensions* (not the row/colnames)
#names(dimnames(dat)) <- c("predicted", "observed")

dat
```

```{r}
# naive Bayes

naive_bayes_df = data.frame(train_df)
naive_bayes_test_df = data.frame(test_df)

naive_bayes_df <- cbind(train_df[1:4], as.data.frame(lapply(train_df[5:11], as.factor)))

for(column in c("atRisk", "headA", "bodyA", "cough", "runny", "nausea", "diarrhea")) {
 
  naive_bayes_df[,column] = as.factor(naive_bayes_df[,column])
  naive_bayes_test_df[,column] = as.factor(naive_bayes_test_df[,column])
}

naive_bayes_model = naiveBayes(atRisk~., naive_bayes_df)
print(naive_bayes_model)
```


```{r}

# ANN alternative
library("neuralnet")
ann_data <- data.frame(train_df)
ann_test_data <- data.frame(test_df)
ann_test_categorical <- ann_test_data[, 12]
for (i in 2:11) {
  ann_data[, i] <- normalize(ann_data[, i])
  ann_test_data[, i] <- normalize(ann_test_data[, i])
}
set.seed(8)
mod_ann <-
  neuralnet(
    atRisk ~ temp + bpSys + vo2 + throat + headA + bodyA + cough + runny + nausea +
      diarrhea,
    ann_data,
    hidden = 5
  )
pred_ann <- compute(mod_ann, ann_test_data)
prediction_data <- pred_ann$net.result
ind <- prediction_data < 0.5
prediction_data[ind] = 0
prediction_data[!ind] = 1
ann_tab <- table(prediction_data, ann_test_data[, 12])
accuracy(ann_tab)
```


Decision Tree:        Accuracy: 81.651376146789 Precision: 80.719794344473 Recall: 87.4651810584958
Naive Bayes           Accuracy: 79.6636085626911 Precision: 79.1237113402062 Recall: 85.5153203342618
SVM                   Accuracy: 78.8990825688073 Precision: 78.2608695652174 Recall: 85.2367688022284
ANN                   Accuracy: 80.8868501529052 Precision: 81.4516129032258 Recall: 84.4011142061281
KNN                   Accuracy: 80.4281345565749 Precision: 79.6915167095116 Recall: 86.3509749303621


Decision Tree:        Accuracy: 81.651376146789 Precision: 80.719794344473 Recall: 87.4651810584958
Naive Bayes           Accuracy: 79.6636085626911 Precision: 79.1237113402062 Recall: 85.5153203342618
SVM                   Accuracy: 78.8990825688073 Precision: 78.2608695652174 Recall: 85.2367688022284
ANN                   Accuracy: 81.8042813455658 Precision: 82.4324324324324 Recall: 84.958217270195"
KNN                   Accuracy: 80.4281345565749 Precision: 79.6915167095116 Recall: 86.3509749303621




t_nb_md <- train(subset(naive_bayes_df, select=-atRisk), naive_bayes_df$atRisk, method = "nb", trControl = trainControl(method = "cv", number = 10))
importance <- varImp(t_nb_md)
plot(importance)











